---
title: "project"
format: html
editor: visual
---

# BDA project



```{r}
SEED = 42
library(aaltobda)
library(rstan)
library(dplyr)
library(ggplot2)
library(loo)
library(bayesplot)
library(gridExtra)
library(grid)
```
## Introduction

## Data description

Here we should write about the data and preprocessing

```{r}
data = read.csv('Salary.csv')
```


Encoding columns


```{r}
data$Gender = as.integer(factor(data$Gender))
data$Job.Title = as.integer(factor(data$Job.Title))
data$Country = as.integer(factor(data$Country))
data$Race = as.integer(factor(data$Race))
```

Splitting for separate model

```{r}
data1 = data %>%  filter(Gender == 1)
data2 = data %>%  filter(Gender == 2)
```



## Models

### Model 1

Priors:
$$N(110000, 50000)$$
Likelihood:

```{r}
separate_data = list(J=2,
                      N1=length(data1$Gender),
                      N2=length(data2$Gender),
                      K=length(data)-1,
                      x1=data1[2:length(data1)],
                      x2=data2[2:length(data2)],
                      y1=data1[['Gender']],
                      y2=data2[['Gender']],
                      prior=1) 
```

Training:

```{r}
separate_fit <- stan(file = "separate_model_project.stan",
                     data = separate_data,
                     iter = 5000,
                     seed = SEED)

sep_samples <- extract(separate_fit)
```


### Model 2

Priors:

Likelihood:

Training:

## Convergence diagnostics

Rhat

Effective sample size

Divergent transitions

Conclusions

## Posterior predictive checks

## LOO-CV comparison

## Sensitivity analysis

Here we should play with priors

## Discussion

Here we should write about issues and potential improvements

## Conclusion

## Self-reflection

What we've learned and struggled with

## Appendix

Stan code here

## References


